Outline paper - Multiple imputation for statistical disclosure control: Creating synthetic data sets with *mice*

### 2021-10-04 mice_synthesizing/paper.Rmd
update results

Unfortunately, this data set does not differ from the vast majority of collected data sets, in the sense that it suffers from missing data. For simplicity, the data is completed using the default `mice` imputation model for all predictors except `bmi`, which is imputed using a passive imputation model due to the deterministic relation with weight and height ($\text{BMI} = \frac{\text{weight}}{\text{height (in meters)}^2}$) to create a single, fully observed data set:

```{r, include = FALSE}
library(mice)
library(magrittr)
library(tidyverse)
library(furrr)
library(knitr)
library(kableExtra)
```

@@

To induce an appropriate amount of sampling variance, 1000 bootstrap samples of the `boys` data will be synthesized with $m = 5$ imputations. The data will be synthesized using the `CART` imputation method for all predictors except BMI, which will be synthesized using passive imputation according to the model `~I(wgt / (hgt / 100)^2)`. Additional parameters that come with the use of `mice.impute.cart()` are the complexity parameter `cp` and the minimum number of observations in any terminal node `minbucket`, that both constrain the flexibility of the imputation model. The values of the parameters `cp` and `minbucket` ought to adhere to the call for imputation models that are as flexible as possible. Appropriate values for these parameters, as well as the input for the `predictorMatrix`, depend on the data at hand. In the current example, the complexity parameter is specified at `cp = 1e-08` rather than the default value `1e-04`, and the minimum number of observations in each terminal node is set at `minbucket` $= 3$ rather than the default value $5$. By allowing for more complexity in the imputation model, bias in the estimates from the synthetic data set is reduced. Additionally, since the missingness pattern is monotone, the number of iterations can be set to `maxit = 1`.

@@

Obtaining a final point estimate of the parameter of interest $Q$ after imputation is fairly easy and no different from pooling in the case of missing data [@rubin_multiple_1987]. One can calculate the average of the $m$ point estimates $q^{(l)}$
$$
\bar{q}_m = \sum_{l = 1}^m \frac{q^{(l)}}{m}.
$$
Also similarly to the missing data case, variances, and subsequently confidence intervals, should incorporate the increase in variance that is due to imputation [@reiter_partially_inference_2003; @drechsler_synthetic_2011]. Yet, the increase in variance due to imputation differs according to whether missing values are imputed or observed data is overimputed with missing values. Whereas the variance estimate after imputation of missing data needs to account for the fact that a certain amount of information in the data is missing, variance estimation from synthetic data does not suffer from this issue. The adjusted variance estimate that follows from using multiple synthetic data sets only suffers from the fact that a finite number of $m$ synthetic data sets are used to resemble the observed data. Hence, the according variance estimate for synthetic data as developed by @reiter_partially_inference_2003 yields
$$
T = \bar{u}_m + \frac{b_m}{m},
$$
with "between-imputation" variance

$$
b_m = \sum_{l = 1}^m \frac{(q^{(l)} - \bar{q}_m)^2}{(m - 1)}, \\
$$
and "within-imputation" variance
$$
\bar{u}_m = \sum_{l = 1} \frac{u^{(l)}}{m},
$$
with $u^{(l)}$ reflecting the variance estimate in every imputed data set.

## Simulation results
```{r, message = FALSE, warning = FALSE, cache = TRUE, cache.lazy = FALSE}
plan(multisession) # increase speed through futures

@@

nsim <- 10 # use 1000 iterations

@@

@@

synthetic_samples[[234]] %>%
  complete(action = "long", include = TRUE) %>% 
  mutate(Synthetic = ifelse(.imp == 0, "Real data", "Synthetic data")) %>%
  ggplot(mapping = aes(x = gen, fill = Synthetic, group = Synthetic)) +
  geom_bar(aes(y = ..prop..), position = "dodge2") +
  scale_fill_brewer(palette = "Paired") +
  theme_classic()

## Merely illustrative, single data set, plot of bmi by region
synthetic_samples[[678]] %>%
  complete(action = "long", include = TRUE) %>%
  mutate(Synthetic = ifelse(.imp == 0, "Real data", "Synthetic data")) %>%
  ggplot(mapping = aes(x = reg, y = bmi, fill = Synthetic)) +
  geom_boxplot() +
  scale_fill_brewer(palette = "Paired") +
  theme_classic()
```

@@

That said, while the field of data synthetisation originated decades ago, it is more relevant today than ever. In our study we outline a standard research workflow where the goal is to synthesize a complete data set. To simplify the simulation we adopted a bootstrapping scheme to induce sampling variance in order to conform to the combination procedure defined by @reiter_partially_inference_2003. Ideally, one would like to omit the bootstrap from the synthetization to adopt a scheme where the sampled data itself serves as a reference: much like the procedure outlined in @vink2014poolingadopt for incomplete data simulation. The corresponding pooling rules have not been derived yet and the incorporation in data analysis workflows would require proper attention from developers alike.

>>>>

Unfortunately, this data set does not differ from the vast majority of collected data sets, in the sense that it suffers from missing data. For simplicity, the data is completed using the default `mice` imputation model for all predictors except `bmi`, which is passively imputed using its deterministic relation with weight and height.

```{r, include = FALSE}
library(mice)
library(tibble)
library(magrittr)
library(tidyverse)
library(furrr)
library(knitr)
library(kableExtra)
library(psych)
library(purrr)
library(dplyr)
library(knitr)
library(kableExtra)
```

@@

To induce sampling variance, 1000 bootstrap samples of the `boys` data will be synthesized with $m = 5$ imputations. Synthetic values are generated using the `CART` imputation method for all columns, except for `bmi`. The deterministic relation `bmi` which will be synthesized passively based on the synthetic values for `hgt` and `wgt` to preserve the relation in the synthetic data. Additional parameters that come with the use of `mice.impute.cart()` are the complexity parameter `cp` and the minimum number of observations in any terminal node `minbucket`, that both constrain the flexibility of the imputation model. The values of the parameters `cp` and `minbucket` ought to adhere to the call for imputation models that are as flexible as possible. Appropriate values for these parameters, as well as the input for the `predictorMatrix`, depend on the data at hand. In the current example, the complexity parameter is specified at `cp = 1e-08` rather than the default value `1e-04`, and the minimum number of observations in each terminal node is set at `minbucket` $= 3$ rather than the default value $5$. By allowing for more complexity in the imputation model, bias in the estimates from the synthetic data set is reduced. Additionally, since the missingness pattern is monotone, the number of iterations can be set to `maxit = 1`.

@@

Obtaining a final point estimate of the parameter of interest $Q$ after imputation is fairly easy and no different from pooling in the case of missing data [@rubin_multiple_1987]. One can calculate the average of the $m$ point estimates $q^{(l)}$
$$
\bar{q}_m = \sum_{l = 1}^m \frac{q^{(l)}}{m}.
$$
with $l = 1, \dots, m$.

Similarly to the missing data case, variances, and subsequently confidence intervals, should incorporate the increase in variance that is due to imputation [@reiter_partially_inference_2003; @drechsler_synthetic_2011]. Yet, the increase in variance due to imputation differs according to whether missing values are imputed or observed data is overimputed with missing values. Whereas the variance estimate after imputation of missing data needs to account for the fact that a certain amount of information in the data is missing, variance estimation from synthetic data does not suffer from this issue. The adjusted variance estimate that follows from using multiple synthetic data sets only suffers from the fact that a finite number of $m$ synthetic data sets are used to resemble the observed data. Hence, the according variance estimate for synthetic data as developed by @reiter_partially_inference_2003 yields
$$
T = \bar{u}_m + \frac{b_m}{m},
$$
with between variance

$$
b_m = \sum_{l = 1}^m \frac{(q^{(l)} - \bar{q}_m)^2}{(m - 1)}, \\
$$
and sampling variance
$$
\bar{u}_m = \sum_{l = 1} \frac{u^{(l)}}{m},
$$
with $u^{(l)}$ reflecting the variance estimate in the $l$th synthetic data set.

# Simulation results
```{r, message = FALSE, warning = FALSE, cache = TRUE, cache.lazy = FALSE}
plan(multisession) # increase speed through futures

@@

nsim <- 1000 # use 1000 iterations

@@

## Univariate estimates
The univariate descriptives for the original data and the synthetic data can be found in the folowing table. [TABLE REF]
```{r}
true <- psych::describe(data)

synth <- synthetic_samples %>% 
  map(~.x %>% complete("all") %>% 
        map(~.x %>% describe()) %>% Reduce("+", .) / synthetic_samples[[1]]$m) %>% 
  Reduce("+", .) / length(synthetic_samples)
rownames(true) <- paste("original", rownames(true), sep = " ")
rownames(synth) <- paste("synthetic", rownames(synth), sep = " ")

bind_rows(true, setNames(synth, names(true))) %>% 
           arrange(vars) %>% .[, c(2:5, 8:9, 11:12)] %>% round(3)
```

$$
% latex table generated in R 4.1.0 by xtable 1.8-4 package
% Sun Oct  3 23:20:37 2021
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrrrr}
  \hline
 & n & mean & sd & median & min & max & skew & kurtosis \\ 
  \hline
age & 748.00 & 9.16 & 6.89 & 10.50 & 0.04 & 21.18 & -0.03 & -1.56 \\ 
  age\_synth & 748.00 & 9.23 & 6.91 & 10.51 & 0.04 & 20.93 & -0.05 & -1.56 \\ 
  hgt & 748.00 & 131.10 & 46.52 & 145.75 & 50.00 & 198.00 & -0.30 & -1.47 \\ 
  hgt\_synth & 748.00 & 131.59 & 46.72 & 146.09 & 50.71 & 197.47 & -0.30 & -1.47 \\ 
  wgt & 748.00 & 37.12 & 26.03 & 34.55 & 3.14 & 117.40 & 0.38 & -1.03 \\ 
  wgt\_synth & 748.00 & 37.44 & 26.23 & 34.87 & 3.36 & 113.04 & 0.36 & -1.05 \\ 
  bmi & 748.00 & 18.04 & 3.04 & 17.45 & 11.73 & 31.74 & 1.14 & 1.79 \\ 
  bmi\_synth & 748.00 & 18.04 & 3.10 & 17.44 & 11.41 & 32.09 & 1.09 & 1.70 \\ 
  hc & 748.00 & 51.62 & 5.86 & 53.10 & 33.70 & 65.00 & -0.91 & 0.12 \\ 
  hc\_synth & 748.00 & 51.65 & 5.88 & 53.19 & 34.30 & 62.12 & -0.93 & 0.16 \\ 
  gen* & 748.00 & 2.53 & 1.59 & 2.00 & 1.00 & 5.00 & 0.52 & -1.36 \\ 
  gen*\_synth & 748.00 & 2.58 & 1.59 & 2.00 & 1.00 & 5.00 & 0.47 & -1.40 \\ 
  phb* & 748.00 & 2.75 & 1.86 & 2.00 & 1.00 & 6.00 & 0.56 & -1.25 \\ 
  phb*\_synth & 748.00 & 2.80 & 1.88 & 2.00 & 1.00 & 6.00 & 0.52 & -1.30 \\ 
  tv & 748.00 & 8.43 & 8.12 & 3.00 & 1.00 & 25.00 & 0.85 & -0.78 \\ 
  tv\_synth & 748.00 & 8.65 & 8.23 & 3.39 & 1.00 & 25.00 & 0.80 & -0.88 \\ 
  reg* & 748.00 & 3.02 & 1.14 & 3.00 & 1.00 & 5.00 & -0.08 & -0.77 \\ 
  reg*\_synth & 748.00 & 3.03 & 1.13 & 3.00 & 1.00 & 5.00 & -0.07 & -0.73 \\ 
   \hline
\end{tabular}
\end{table}
$$

We see that the synthetic data estimates resemble the true data estimates very well.

@@

synthetic_samples[[2]] %>%
  complete(action = "long", include = TRUE) %>% 
  mutate(Synthetic = ifelse(.imp == 0, "Real data", "Synthetic data")) %>%
  ggplot(mapping = aes(x = gen, fill = Synthetic, group = Synthetic)) +
  geom_bar(aes(y = ..prop..), position = "dodge2") +
  scale_fill_brewer(palette = "Paired") +
  theme_classic()

## Merely illustrative, single data set, plot of bmi by region
synthetic_samples[[3]] %>%
  complete(action = "long", include = TRUE) %>%
  mutate(Synthetic = ifelse(.imp == 0, "Real data", "Synthetic data")) %>%
  ggplot(mapping = aes(x = reg, y = bmi, fill = Synthetic)) +
  geom_boxplot() +
  scale_fill_brewer(palette = "Paired") +
  theme_classic()
```

@@

That said, while the field of data synthetisation originated decades ago, it is more relevant today than ever. In our study we outline a standard research workflow where the goal is to synthesize a complete data set. To simplify the simulation we adopted a bootstrapping scheme to induce sampling variance in order to conform to the combination procedure defined by @reiter_partially_inference_2003. Ideally, one would like to omit the bootstrap from the synthetization to adopt a scheme where the sampled data itself serves as a reference: much like the procedure outlined in @vink2014pooling for incomplete data simulation. The corresponding pooling rules have not been derived yet and the incorporation in data analysis workflows would require proper attention from developers alike.
